---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

# analysiskmeans

<!-- badges: start -->
[![R-CMD-check](https://github.com/nsmodi/analysiskmeans/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/nsmodi/analysiskmeans/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/nsmodi/analysiskmeans/actions/workflows/pkgdown.yaml/badge.svg)](https://github.com/nsmodi/analysiskmeans/actions/workflows/pkgdown.yaml)
<!-- badges: end -->

The goal of analysiskmeans is to perform a K Means Clustering on a Single Cell Experiment data object!

## Installation

You can install the development version of analysiskmeans from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("nsmodi/analysiskmeans")
```
Or you can also ...

``` r
# library(devtools)
# remotes::install_github("nsmodi/analysiskmeans")
```
##Functions this package provides!

## Example: data_config()

This is a basic example of how to use the data_config function to set up your SingleCellExperiment data so that it is usable for functionality in this package!

```{r example, echo=F}
  library(analysiskmeans)
  utils::data(example_sce, package="analysiskmeans")
  sce <- example_sce
  results <- data_config(sce)
  sce <- results$sce
```
## Example: top_x_genes()

Next, we can take a look creating a subset of our data including a particular number of top selected genes! Keep in mind that if you do not fill in a parameter value for n_top, 50 will be the default value. A small portion of this subset is shown below. 

```{r example2}
  utils::data(example_sce, package="analysiskmeans")
  sce <- example_sce
  results <- data_config(sce)
  sce <- results$sce
  mat_norm <- top_x_genes(sce, n_top = 50)
mat_norm[1:5,1:5]
```
 
## Example: computepca()

We can now compute a PCA analysis as a part of doing Kmeans. Within this, we will need to pass the mat_norm object we just created into the computepca() function. A few outputs of the pca function are shown below. 

```{r pca}
  utils::data(example_sce, package="analysiskmeans")
  sce1 <- example_sce
  results1 <- data_config(sce1)
  sce1 <- results1$sce
  mat_norm1 <- top_x_genes(sce1, n_top = 50)
  pca1 <- computepca(mat_norm1)
  pca1$sdev[1:5]
```

## Example: kmeans()

The kmeans function conducts clustering in an iterative fashion, going from inputted parameters of minimum k values to maximum k values. The km_list object contains a large amount of information about the clustering. The metrics object contains wss and average silhouette scores with corresponding to each k value in our range. An example of what metrics looks like is shown below.

```{r kmeans}
  utils::data(example_sce, package="analysiskmeans")
  sce <- example_sce
  results <- data_config(sce)
  sce <- results$sce
  mat_norm <- top_x_genes(sce, n_top = 50)
  pca <- computepca(mat_norm)
  max_k <- 10
  min_k <-5
  outputs <- k_means(min_k = min_k, max_k=max_k, pca = pca)
  metrics<-outputs$metrics
  km_list <- outputs$km_list
  metrics
```

## Example: cluster_plot() 

Our cluster_plot() function takes outputs from the kmeans function and plots the data based on principal components. An example plot is shown above, and each cluster is differentiated by color. We plot the following based on an example selected k value. In this case we have chosen a value of 8 as an example.

```{r cluster}
plot <- cluster_plot(selected_k=8, km_list, pca, results$cell_type)
```
## Using analysiskmeans in CLI

The analysiskmeans package can also be accessed by the command line. A few example of prompts are shown below!

To get started, if you are looking for the different commands you can use with this package, use the --help command. 

```
analysiskmeans --help
```
If you are looking to retrieve an elbow plot for various Kmeans clustering iterations, the *elbow* command is the one to use! Here are the following methods to use with the elbow command: 

```
Options:
  -c, --counts <COUNTS>       Path to counts matrix (TSV/CSV, genes x samples)
                              [default: ""] [type: string]
  -gm, --genemeta <GENEMETA>  Path to sample gene metadata (TSV/CSV)
                              [default: ""] [type: string]
  -cm, --cellmeta <CELLMETA>  Path to sample cell metadata (TSV/CSV)
                              [default: ""] [type: string]
  -o, --output <OUTPUT>       Output directory [default: ""] [type: string]
  -FALSE, --n-top <N-TOP>     Number of top variable genes
                              [default: 50] [type: integer]
  -mn, --min-k <MIN-K>        Minimum K Value [default: 5] [type: integer]
  -mx, --max-k <MAX-K>        Number of top variable genes
                              [default: 10] [type: integer]
```
If you are looking to retrieve clustering kmeans plots for various Kmeans clustering iterations, the *cluster* command is the one to use! Here are the following methods to use with the cluster command: 

```
Options:
  -c, --counts <COUNTS>           Path to counts matrix (TSV/CSV, genes x
                                  samples)
                                  [default: ""] [type: string]
  -gm, --genemeta <GENEMETA>      Path to sample gene metadata (TSV/CSV)
                                  [default: ""] [type: string]
  -cm, --cellmeta <CELLMETA>      Path to sample cell metadata (TSV/CSV)
                                  [default: ""] [type: string]
  -o, --output <OUTPUT>           Output directory [default: ""] [type: string]
  -FALSE, --n-top <N-TOP>         Number of top variable genes
                                  [default: 50] [type: integer]
  -mn, --min-k <MIN-K>            Minimum K Value [default: 5] [type: integer]
  -mx, --max-k <MAX-K>            Number of top variable genes
                                  [default: 10] [type: integer]
  -mx, --selected-k <SELECTED-K>  A Selected K Value for Plotting
                                  [default: 7] [type: integer]
```

Let's get started with a few examples!


Firstly, read_data_file is a helper function to load in a dataset in a csv/tsv format
```{r cli}
library(SingleCellExperiment)
read_data_file <- function(path) {
  ext <- tolower(tools::file_ext(path))
  if (ext == "csv") {
    utils::read.csv(path, row.names = 1, check.names = FALSE)
  } else {
    utils::read.table(path, sep = "\t", header = TRUE, row.names = 1,
                      check.names = FALSE)
  }
}
```

The following example starts by constructing input files!

```{r cli2}
set.seed(42)
num_genes <- 100
num_cells <- 20 #Used to be 8
raw_counts <- as.integer(rexp(num_genes*num_cells, rate = 0.1))
raw_counts <- matrix(raw_counts, nrow = num_genes, ncol = num_cells)

gene_metadata <- data.frame(
  name = paste("Gene", 1:num_genes, sep = "_"),
  length = as.integer(rnorm(num_genes, mean = 10000, sd = 500))
)

cell_metadata <- data.frame(
  names = paste("Cell", 1:num_cells, sep = "_"),
  batch = rep(1:2, each = num_cells/2),
  label = rep(c("xylem", "phloem"), times = num_cells/2)
)
```


We can continue by writing the inputs into files in a tests/cli folder!
```{r cli3}
#reading the files and beginning the CLI simulation
counts_df <- read_data_file(paste(getwd(),"/tests/cli/counts_df.csv", sep = ""))
genemeta_df <- read_data_file(paste(getwd(),"/tests/cli/genemeta_df.csv", sep = ""))
cellmeta_df <- read_data_file(paste(getwd(),"/tests/cli/cellmeta_df.csv", sep = ""))
```

Next, we can make an SingleCellExperiment object with the dataframes from the files!

```{r cli4}

sce1 <- SingleCellExperiment(
  assays = list(counts = as.matrix(counts_df)),
  rowData = genemeta_df,
  colData = cellmeta_df
)
```

After the file is fetched, and the object is saved, we can run our functions and start conducting the analysis! We can start by using the elbow plot function! We are also saving our elbow plot in an image in our repository. 

```{r cli5}
results <- data_config(sce1)
sce1 <- results$sce
mat_norm <- top_x_genes(sce1, n_top = 50)
pca <- computepca(mat_norm)
max_k <- 10
min_k = 5
outputs <- k_means(min_k = min_k, max_k=max_k, pca = pca)
metrics<-outputs$metrics
grDevices::png(paste(getwd(),"/tests/cli/plots/elbowplot.png", sep = ""), width = 8, height = 6, units = "in", res = 300)
elbow_plot(metrics)
message("Saved!")
```






